---
title: "Design of surveys"
output: learnr::tutorial
runtime: shiny_prerendered
description: Questions concerning aerial survey of marine mammals near St Andrews and terrestrial point transect surveys for birds also near St Andrews.
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


<img src=https://images.unsplash.com/photo-1526485796220-1c4ad338f85f?ixid=MXwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHw%3D&ixlib=rb-1.2.1&auto=format&fit=crop&w=1351&q=80 width=400 height=200 style="float:right">

## Exercise 6 -- Survey design

Output provided by the `dssd` package is extensive; the following is intended to help you find the most important aspects of that output so you can effectively assess the merits of the survey designs you produce. The following questions are intended to have you pick through the output and *think* about what some of those values are telling you.  The questions also ask you to do some "sense checking" to ensure the software is performing the way you expect (never trust software to do the right thing).

```{r  message=FALSE, warning=FALSE, echo=FALSE}
library(dssd)
shapefile.name <- system.file("extdata", "StAndrew.shp", package = "dssd")
region <- make.region(region.name = "St Andrews Bay",
                      units = "m",
                      shape = shapefile.name)
cover <- make.coverage(region, n.grid.points = 2000)
design.space500 <- make.design(region = region,
                      transect.type = "line",
                      design = "systematic",
                      spacing = 5000,
                      design.angle = 90,
                      edge.protocol = "minus",
                      truncation = 2000,
                      coverage.grid = cover)
design.space500 <- run.coverage(design.space500, reps = 100, quiet=TRUE)
print(design.space500)
```


### Answer these questions by examining at the output above

Use some common sense about the parallel aerial survey you have designed.  If the design specification is to place transects 5000m (5km) apart and further if truncation distance is 2000m (on either side of the aircraft), employ some simple geometry to assess whether `dssd` is doing the right thing.


```{r spatialcover}
question_numeric("What should be the percentage of the study area covered by survey effort? (to the nearest 5 percent)",
   tolerance=5, allow_retry = TRUE, 
   answer=80, correct=TRUE)
```

The design problem you were to solve treated the fuel capacity of the aircraft as the constraint around which you needed to operate.  Remember, the aircraft chosen could only travel 250km without refueling.  That constraint lead you to produce a survey in which the parallel transects were spaced at 5km.

However, now take a broader view of the problem from a statistical rather than a logistical perspective.  Look in the output provided to determine the number of replicate transects resulting from spacing the transects 5km apart and orienting the transects east-to-west.


```{r reps}
quiz(
question_numeric("How many replicate transects result from this design?",
   tolerance=1, allow_retry = TRUE, 
   answer(8, correct=TRUE)
   ),
question_radio("Is that sufficient replication (based upon what you learned from the variance discussion Thursday)?",
   answer("yes"),
   answer("no", correct=TRUE)
   ),
question_checkmark("What possible solutions (more than one answer) might there be to this lack of replication?",
  answer("find another study area"),
  answer("hire an aircraft with greater fuel capacity", correct=TRUE),
  answer("use segmented transects", correct=TRUE),
  answer("shorten the truncation distance")
  )
)
```

## Comparing parallel and zigzag design

Design-based inference rests upon the premise every location in the study area has the same chance of being sampled as every other location.  In other words, coverage scores is uniform.  We use the heat maps created by plotting design objects to visually assess this uniformity.

```{r echo=FALSE}
design.zz.4500 <- make.design(region = region,
                      transect.type = "line",
                      design = "eszigzag",
                      spacing = 4500,
                      design.angle = 0,
                      edge.protocol = "minus",
                      bounding.shape = "convex.hull",
                      truncation = 2000,
                      coverage.grid = cover)
design.zz.4500 <- run.coverage(design.zz.4500, reps = 100, quiet=TRUE)
print(design.zz.4500)
```

```{r ZIG}
quiz(
question_radio("What numerical output might provide a more objective relative assessment of uniformity between designs?",
   answer("mean line length"),
   answer("sd (standard deviation) of trackline length"),
   answer("median coverage score"),
   answer("sd (standard deviation) of coverage score", correct=TRUE),
   ),
question_radio("Using the metric you chose in the previous question, which (parallel or zigzag) appears to have the more uniform coverage score?",
   answer("parallel", correct=TRUE),
   answer("zigzag")
   )
)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```
