---
title: "Detection functions"
output: learnr::tutorial
runtime: shiny_prerendered
description: Re-visit duck nest data, exploring model output to locate important printed quantities.
---

```{r setup, include=FALSE}
library(learnr)
library(magrittr)
knitr::opts_chunk$set(echo = FALSE)
```

<img src=https://images.unsplash.com/photo-1593594820228-a05cc04bc433?ixid=MXwxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHw%3D&ixlib=rb-1.2.1&auto=format&fit=crop&w=1350&q=80  width=400 height=200 style="float:right">

## Practical 2 -- Duck nest analysis

This has been your first experience using the `Distance` package.  There are only a handful of functions you need to successfully complete a distance sampling analysis.  This practical gave you experience with these functions.  The data are familiar to you because they the same data you used when trying to fit a detection function by hand in Exercise 1.  This exercise lets the computer do the work.  Compare the estimate of duck nest density produced by `Distance` with the estimate you manually produced.

You should have already executed code as described in Practical 2.  I have duplicated some of the code and output here to reduce your need to switch between this tutorial and your previous work.

```{r message=FALSE,  echo=FALSE, comment=NA, R.options = list(width =90)}
library(Distance, quietly = TRUE)
data(ducknest)
duckunits <- convert_units("meter", "kilometer", "square kilometer")
halfnorm <- ds(data=ducknest, key="hn", convert_units = duckunits)
summary(halfnorm)
```

```{r nests}
quiz(
  question("For the half normal detection function, what was the estimated value of $\\hat{P}_a$?",
   answer("0.869", correct=TRUE),
   answer("0.039"),
   answer("614.25"),
   answer("0.933")
   ),
  question("Given your estimate of $\\hat{P}_a$, what does it mean?",
    answer("distance to which nests were detected"),
    answer("area under the curve"),
    answer("proportion of nests detected out to 2.4m", correct=TRUE),
    answer("number of iterations needed to estimate detection function parameters")
  )
)
```

## Convert density to abundance

The field `area` was set to 0 in the `ducknest` data frame; however the practical description notes that the refuge is 47.7 $km^2$ in size.  Use the code block below to complete the abundance estimate by computing the number of duck nests on the refuge based upon your estimate of $\hat{D}$.

Remember the density estimate from the half normal key function model was **`r round(halfnorm$dht$individuals$D$Estimate,2)`**.

Replace the blanks in  the following code to produce an estimate of the number of nests on the refuge.

```{r abundance, exercise=TRUE, exercise.eval=FALSE, exercise.blanks=TRUE}
dhat <- ___
area <- ___
est.num.nests <- ___
print(round(est.num.nests))
```

```{r abundance-hint-1}
dhat <- 49.7
```

```{r abundance-hint-2}
dhat <- 49.7
area <- 47.7
```

```{r abundance-hint-3}
dhat <- 49.7
area <- 47.7
est.num.nests <- dhat * area
```


```{r numnests}
question_numeric("To the nearest 10, what is the estimated total number of nests on the refuge?",
  tolerance=10, allow_retry = TRUE, 
  answer(2370, correct=TRUE)
)
```


## Does a model fit the data?

Although we have not yet discussed how to assess fit of our models to data, Practical 2 did introduce you to the function `gof_ds`.  The code below applies the goodness of fit function to the half normal detection function fitted to the duck nest data. You are to interpret the output.

```{r gof, comment=NA}
gof_ds(halfnorm)
```


```{r gof-quest}
quiz(
  question_checkbox("How should the plot be interpreted? (select all that apply)",
    answer("departures from good fit are indicated by departures from the underlying 45° line", correct=TRUE),
    answer("model that fits has all points near the 45° line", correct=TRUE),
    answer("this plot has nothing to do with model fit"),
    answer("a rapid rise in the graphed points implies good model fit")
    ),
  question("The numerical output from `gof_ds` is from the Cramer-von Mises test.   Interpret the output.",
    answer("P-value is <1, therefore the model fit is poor"),
    answer("The test statistic is smaller than the P-value, therefore the fit is adequate"),
    answer("The test statistic is close to 0, therefore the fit is poor"),
    answer("P-value is close to one, indicating that there’s no reason to suspect the fit is inadequate", correct=TRUE)
    )
)

```

## Density estimates from different models

```{r echo=FALSE, message=FALSE}
nest.hn <- ds(ducknest, key="hn", adjustment=NULL,
              convert_units=duckunits)
nest.uf.cos <- ds(ducknest, key="unif", adjustment="cos",
                  convert_units=duckunits)
nest.hr.herm <- ds(ducknest, key="hr", adjustment="herm", 
                  convert_units=duckunits)
```

```{r table, echo=FALSE}
nest.tab <- data.frame(
                       DetectionFunction=c("Half-normal, no adjustments",
                                           "Uniform, cosine adjustments",
                                           "Hazard rate, no adjustments "),
                       Density=rep(NA,3), LowerCI=rep(NA,3), UpperCI=rep(NA,3))

get.results.f <- function(fit.model) {   
  return(c(D=fit.model$dht$individuals$D$Estimate,
         lCL=fit.model$dht$individuals$D$lcl,
         uCL=fit.model$dht$individuals$D$ucl))
}
nest.tab[1,2:4] <- get.results.f(nest.hn)
nest.tab[2,2:4] <- get.results.f(nest.uf.cos)
nest.tab[3,2:4] <- get.results.f(nest.hr.herm)
knitr::kable(nest.tab, 
             caption="Density estimates and confidence intervals for three fitted models.", 
             digits = 1) %>%
    kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

The preceding code fit all three key function models to the duck nest data.  To reinforce the idea that similar results are produced *(for good data)* from the different key functions, write a few lines of code to explore this, using **R** as a calculator.  Value entered to one decimal place.


```{r dendiff, exercise=TRUE, exercise.eval=FALSE, exercise.blanks="___+"}
smallest_est <- ___
largest_est <- ___
difference <- largest_est - smallest_est
relative_difference <- difference / smallest_est
print(round(relative_difference,3))
```

```{r dendiff-hint-1}
smallest_est <- 48.6
```

```{r dendiff-hint-2}
smallest_est <- 48.6
largest_est <- 51
```


From the calculation above,

```{r diffquest}
question_numeric("To the nearest **percent**, what is the relative difference between the largest and smallest density estimate you just computed?",
  min = 0, max = 10, step = 1, allow_retry = TRUE, incorrect="The answer is in percent, not proportion.",
  answer(5, correct=TRUE)
)
```
